Manuscript Figures

### 2018-05-25 Manuscript Figures.Rmd
Cleaning up new report for just reporting manuscript figures
This update leave the original subgenome report broken, but makes the manuscript figures report functional.  Makes use of updated MaizeMap/MaizeGO/MaizeOmics versions in the figures, cleans up a lot of the unnecessary analysis.

# All EV-Codes, GOSlim only
grid.newpage()
g <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm), 
                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm)) %>% 
              nrow(), 
  category = c("Sub1", "Sub2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim only)")

@@

#EXP only, GOSlim only
grid.newpage()
g <- draw.pairwise.venn(
  area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
  area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
  cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% select(goTerm), 
                         goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm)
                         ) %>% nrow(), 
  category = c("Sub1", "Sub2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim only)")

@@

```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
plot <- graphGenePairDominanceByExperiment(maize.expression.sample.avg, homeologs.pairs, 4, FALSE)
plot(plot)
cap <- "Fig 9.1:"
```

@@

```{r subgenome_protein_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
factor <- 4
plot <- graphGenePairDominanceByExperiment(maize.protein.abundance.sample.avg, homeologs.pairs, factor, FALSE) + 
  labs(title = paste0("Which homeolog has greater protein abundance? (factor=",factor,")"))
plot(plot)
cap <- "Fig 9.2:"
```

@@

```{r subgenome_expression_dominance_across_factors, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=12, fig.align='center', fig.cap = cap}
plot <- graphGenePairDominanceByExperimentAcrossFactors(maize.expression.sample.avg, homeologs.pairs, FALSE)
plot(plot)
cap <- "Fig 9.3:"
```

@@

Looking at a few individual cases (* these are calculated as maize1 > maize2, since maize2 > maize1 makes a negative fold change):

```{r subgenome_expression_top_ten, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
gene.foldchange <- homeologFoldChanges(maize.expression.sample.avg, homeologs.pairs)
gene.foldchange$Maize1_url <- apply(gene.foldchange[,1], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
gene.foldchange$Maize2_url <- apply(gene.foldchange[,2], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
gene.foldchange[1:10,] %>% 
  select(Maize1_url, Maize2_url, sample, Value_maize1, Value_maize2, foldChange) %>%
  knitr::kable(caption = "Top ten differentially expressed gene pairs (in log2)", align = "llrrr")

protein.foldchange <- homeologFoldChanges(maize.protein.abundance.sample.avg, homeologs.pairs)
protein.foldchange$Maize1_url <- apply(protein.foldchange[,1], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
protein.foldchange$Maize2_url <- apply(protein.foldchange[,2], 2, function(x) { wrapInHREF(getGenePageURL(x), x) })
protein.foldchange[1:10,] %>% 
  select(Maize1_url, Maize2_url, sample, Value_maize1, Value_maize2, foldChange) %>%
  knitr::kable(caption = "Top ten differentially abundant protein pairs (in log2)", align = "llrrr")

gene.foldchange[1:50,] %>%
  inner_join(protein.foldchange[1:50,], by = c("Maize1", "Maize2", "sample")) %>%
  rename(foldChange_expr=foldChange.x, foldChange_prot=foldChange.y, Maize1_url=Maize1_url.x, Maize2_url=Maize2_url.x) %>%
  select(Maize1_url, Maize2_url, sample, foldChange_expr, foldChange_prot) %>%
  knitr::kable(caption = "Gene pairs in the top 50 for both differential gene expression and differential protein abundance (in log2).", align = "llrrrr")
```

Look at the top 10 pairs with highest isoform count diffs (foldChange in log2)

```{r subgenome_isoforms_top_ten, echo=FALSE, warning=FALSE, message=FALSE, results='asis'}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(geneTranscript.counts, by=c("Maize1"="gene")) %>%
  inner_join(geneTranscript.counts, by=c("Maize2"="gene"))
names(data)[3] <- "Value_maize1"
names(data)[4] <- "Value_maize2"
data <- subset(data, !is.na(data$Value_maize1) & !is.na(data$Value_maize2))

nSingleExonGenes.total <- nrow(geneTranscript.counts[geneTranscript.counts$n == 1,])
nSingleExonGenes.inSet <- unique(c(data$Maize1[data$Value_maize1 == 1], data$Maize2[data$Value_maize2 == 1]))

data <- subset(data, data$Value_maize1 > 1 & data$Value_maize2 > 1) ## Ignore single-exon genes
data$foldChange <- log2(data$Value_maize1) - log2(data$Value_maize2)
data <- arrange(data, desc(foldChange))
data$Maize1_url <- apply(data[,1], 2, function(x) { wrapInHREF(getGBrowsePageURL(x), x) })
data$Maize2_url <- apply(data[,2], 2, function(x) { wrapInHREF(getGBrowsePageURL(x), x) })
data[1:10,] %>%
  select(Maize1_url, Maize2_url, Value_maize1, Value_maize2, foldChange) %>%
  knitr::kable(caption = "Top 10 pairs with highest isoform count diffs (foldChange in log2)", align = "llrrr")
```

@@

```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

@@

```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID")) %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

@@

```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center', fig.cap = cap}
## Start with the paired expression data
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

@@

```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
## Start with the paired expression data
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

@@

# 11 Data Exploration

```{r include=FALSE}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.expression.sample.avg, by=c("Maize1"="geneID")) %>%
  inner_join(maize.expression.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.protein.abundance.sample.avg, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

data$foldChange_expr <- log2(data$FPKM_maize1) - log2(data$FPKM_maize2)
data$foldChange_abun <- log2(data$dNSAF_maize1) - log2(data$dNSAF_maize2)
```

```{r echo=FALSE, out.width='100%'}
knitr::include_graphics('./data-raw/Images/High expressing genes tend to vary.png')
```

```{r echo=FALSE, out.width='100%'}
knitr::include_graphics('./data-raw/Images/Low Expressing Genes tend to stay low.png')
```

```{r echo=FALSE, out.width='100%'}
knitr::include_graphics('./data-raw/Images/Pollen is different.png')
```

```{r echo=FALSE, out.width='100%'}
knitr::include_graphics('./data-raw/Images/High-low variation.png')
```

Red Square = [Zm00001d020592](https://www.maizegdb.org/gene_center/gene/Zm00001d020592) Pericarp/Aleurone 27 DAP vs. [Zm00001d005793](https://www.maizegdb.org/gene_center/gene/Zm00001d005793)

Yellow Square = [Zm00001d003188](https://www.maizegdb.org/gene_center/gene/Zm00001d003188) Pericarp/Aleurone 27 DAP vs. [Zm00001d025753](https://www.maizegdb.org/gene_center/gene/Zm00001d025753)

Yellow X = Zm00001d039040 Mature Leaf 8

Pink Cross = Zm00001d015385 Mature Leaf 8

```{r echo=FALSE, out.width='100%'}
knitr::include_graphics('./data-raw/Images/Top fold-change genes.png')
```

>>>>

# # All EV-Codes, GOSlim only
# grid.newpage()
# g <- draw.pairwise.venn(
#   area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
#   area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm) %>% distinct() %>% nrow(),
#   cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm), 
#                          goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% select(goTerm)) %>% 
#               nrow(), 
#   category = c("Sub1", "Sub2"), 
#   lty = rep("blank", 2), 
#   fill = c("light blue", "green"), 
#   alpha = rep(0.5, 2), 
#   cat.pos = c(0, 180), 
#   scaled = FALSE,
#   euler.d = FALSE, 
#   sep.dist = 0.3, 
#   rotation.degree = 0,
#   ind = FALSE
# )
# grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(All EV-Codes, GOSlim only)")

@@

# #EXP only, GOSlim only
# grid.newpage()
# g <- draw.pairwise.venn(
#   area1 = goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
#   area2 = goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm) %>% distinct() %>% nrow(),
#   cross.area = intersect(goAnnotations.sub1 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm"))%>% filter(evCode=="EXP") %>% select(goTerm), 
#                          goAnnotations.sub2 %>% inner_join(go.goSlim.plant, by=c(goTerm = "GOSlimTerm")) %>% filter(evCode=="EXP") %>% select(goTerm)
#                          ) %>% nrow(), 
#   category = c("Sub1", "Sub2"), 
#   lty = rep("blank", 2), 
#   fill = c("light blue", "green"), 
#   alpha = rep(0.5, 2), 
#   cat.pos = c(0, 180), 
#   scaled = FALSE,
#   euler.d = FALSE, 
#   sep.dist = 0.3, 
#   rotation.degree = 0,
#   ind = FALSE
# )
# grid.arrange(gTree(children=g), top="GO Term by Subgenome\n(EV-EXP, GOSlim only)")

@@

```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
plot <- graphGenePairDominanceByExperiment(maize.walley.v4mapped.expression.data, homeologs.pairs, 4, FALSE)
plot(plot)
cap <- "Fig 9.1:"
```

@@

```{r subgenome_protein_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=6, fig.align='center', fig.cap = cap}
factor <- 4
maize.walley.abundance.v4.data <- 
  maize.walley.abundance.v4 %>%
  select(geneID, sample, dNSAF_avg)
plot <- graphGenePairDominanceByExperiment(maize.walley.abundance.v4.data, homeologs.pairs, factor, FALSE) + 
  labs(title = paste0("Which homeolog has greater protein abundance? (factor=",factor,")"))
plot(plot)
cap <- "Fig 9.2:"
```

@@

```{r subgenome_expression_dominance_across_factors, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=12, fig.align='center', fig.cap = cap}
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
plot <- graphGenePairDominanceByExperimentAcrossFactors(maize.walley.v4mapped.expression.data, homeologs.pairs, FALSE)
plot(plot)
cap <- "Fig 9.3:"
```

@@

@@

```{r subgenome_expression_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

@@

```{r subgenome_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[4] <- "Value_maize1"
names(data)[5] <- "Value_maize2"
# data <- subset(data, sample %in% c("Silk", "Embryo 20 DAP", "Leaf Zone 2 (Stomatal)"))

@@

```{r subgenome_expression_protein_foldchange_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=6, fig.height=5, fig.align='center', fig.cap = cap}
## Start with the paired expression data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

@@

```{r subgenome_expression_protein_scatter, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=8, fig.align='center', fig.cap = cap}
## Start with the paired expression data
maize.walley.v4mapped.expression.data <- 
  maize.walley.v4mapped.expression %>%
  select(geneID, sample, FPKM_avg)
data <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>%
  select(Maize1, Maize2) %>%
  distinct() %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize1"="geneID")) %>%
  inner_join(maize.walley.v4mapped.expression.data, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(FPKM_avg.x) & !is.na(FPKM_avg.y))
names(data)[4] <- "FPKM_maize1"
names(data)[5] <- "FPKM_maize2"

## Add paired abundance data
data <-
  data %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize1"="geneID", "sample"="sample")) %>%
  inner_join(maize.walley.abundance.v4, by=c("Maize2"="geneID", "sample"="sample")) %>%
  subset(!is.na(dNSAF_avg.x) & !is.na(dNSAF_avg.y))
names(data)[6] <- "dNSAF_maize1"
names(data)[7] <- "dNSAF_maize2"

@@
